A Neural Network Alternative to Convolutive Audio Models for Source Separation

### 2017-06-07 paper/paper.tex
Add files via upload

\begin{align}
    \text{$1^{\text{st}}$ layer:~}\mathbf{H}(i,t) &= \sum_{i=1}^{r} \sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k) \\
    \text{$2^{\text{nd}}$ layer:~}\hat{\mathbf{X}}(f,t) &= \sum_{i=1}^{r} \sum_{k=0}^{T-1} \mathbf{W}_{i}(k,f)\cdot\mathbf{H}(i,t-k)
    \label{eq:nmfcnncnn}
\end{align}
subject to non-negativity of $\mathbf{W}_{i}$ and $\mathbf{H}\geq0$. Here, we assume that the convolutional filters~$\mathbf{W},~\mathbf{W}^{\ddagger}$ have a size of $M\times T$ where, $T$ represents the depth of the convolution. In this representation, $\mathbf{W}_{i}$ and $\mathbf{H}$ correspond to the $i^{\text{th}}$ basis matrix and the activation matrix respectively. The filters of the first convolutional layer act as inverse filters in defining the auto-encoder. In the remainder of this section, we will refer to the first convolutional layer as the encoder that estimates a code from the input representation. The second convolutional layer generates an approximation of the input from the code and will be referred to as the decoder. We can simplify the non-negativity constraints by incorporating a non-linearity into the definitions of the encoder and the decoder. Thus,

\begin{align}
    \text{$1^{\text{st}}$ layer:~}\mathbf{H}(i,t) &= g\left(\sum_{i=1}^{r} \sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k)\right) \\
    \text{$2^{\text{nd}}$ layer:~}\hat{\mathbf{X}}(f,t) &= g\left(\sum_{i=1}^{r} \sum_{k=0}^{T-1} \mathbf{W}_{i}(k,f)\cdot\mathbf{H}(i,t-k)\right)
    \label{eq:nmfcnncnn}
\end{align}
Here, the non-linearity $g(.):\mathbb{R}\rightarrow \mathbb{R^{\geq0}}$ applies an element-wise non-negativity constraint and ensures that the activation matrix and the reconstruction are non-negative. We now note a couple of key points about the convolutional auto-encoder. (i) The output of the encoder gives the latent representation of the decomposition. (ii) The weights of the decoder act as the basis vectors of the decomposition. (iii) We do not explicitly apply non-negativity constraints on the weights. Thus, the basis matrices (decoder filters) can assume negative values.

@@

\subsection{Practical Considerations}
\label{subsec:practical}
Having developed the convolutional auto-encoder equivalent to convolutive NMF, we can now begin to understand the nature of the basis and activation matrices learned by the network. To do so, we train the convolutive auto-encoder defined by~(\ref{eq:nmfcnncnn}) on a simple toy example as shown in figure~\ref{}(**Figure reference**)

@@

\section{Supervised Source Separation}
\label{sec:ss}
Describe the separation procedure.

>>>>

\begin{align}
    \text{$1^{\text{st}}$ layer:~}\mathbf{H}(i,t) &= \sum_{i=1}^{r} \sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k) \\
    \text{$2^{\text{nd}}$ layer:~}\hat{\mathbf{X}}(f,t) &= \sum_{i=1}^{r} \sum_{k=0}^{T-1} \mathbf{W}_{i}(k,f)\cdot\mathbf{H}(i,t-k)
    \label{eq:nmfcnncnn}
\end{align}
subject to non-negativity of $\mathbf{W}_{i}$ and $\mathbf{H}$. Here, we assume that the convolutional filters~$\mathbf{W},~\mathbf{W}^{\ddagger}$ have a size of $M\times T$ where, $T$ represents the depth of the convolution. In this representation, $\mathbf{W}_{i}$ and $\mathbf{H}$ correspond to the $i^{\text{th}}$ basis matrix and the activation matrix respectively. The filters of the first convolutional layer act as inverse filters in defining the auto-encoder. In the remainder of this section, we will refer to the first convolutional layer as the encoder that estimates a code from the input representation. The second convolutional layer generates an approximation of the input from the code and will be referred to as the decoder. We can simplify the non-negativity constraints by incorporating a non-linearity into the definitions of the encoder and the decoder. Thus,

\begin{align}
    \text{$1^{\text{st}}$ layer:~}\mathbf{H}(i,t) &= g\left(\sum_{i=1}^{r} \sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k)\right) \\
    \text{$2^{\text{nd}}$ layer:~}\hat{\mathbf{X}}(f,t) &= g\left(\sum_{i=1}^{r} \sum_{k=0}^{T-1} \mathbf{W}_{i}(k,f)\cdot\mathbf{H}(i,t-k)\right)
    \label{eq:nmfcnncnn}
\end{align}
Here, the non-linearity $g(.):\mathbb{R}\rightarrow \mathbb{R}^{\geq0}$ applies an element-wise non-negativity constraint and ensures that the activation matrix and the reconstruction are non-negative. We now note a couple of key points about the convolutional auto-encoder. (i) The output of the encoder gives the latent representation of the decomposition. (ii) The weights of the decoder act as the basis vectors of the decomposition. (iii) We do not explicitly apply non-negativity constraints on the weights. Thus, the basis matrices (decoder filters) can assume negative values.

@@

\subsection{Practical Considerations}
\label{subsec:practical}
Having developed the convolutional auto-encoder equivalent to convolutive NMF, we can now begin to understand the nature of the basis and activation matrices learned by the network. To do so, we train the convolutive auto-encoder defined by~(\ref{eq:nmfcnncnn}) on a simple toy example as shown in figure~\ref{}(**Figure reference**)7

@@

\section{Supervised Source Separation}
\label{sec:ss}
The problem of supervised source separation is solved as a two-step procedure~\cite{}(**Paris supervised and semi-supervised**). The fist step of the procedure is to learn suitable models for a given source. We refer to this step as the training step. In the second step, we use these models to explain the contribution of the source in an unknown mixture. In section~\ref{sec:conv-nmf}, we have developed the auto-encoder architecture to learn suitable convolutive models for a given source. We now turn our attention to the problem of using the models for separating the source in an unknown mixture.

The previous approach to source separation~\cite{}(**ICASSP paper**) involves estimating the latent representation of the bases, which captures the contribution of the bases to each frame of the mixture spectrogram. This does not utilize the encoder component of the auto-encoder for the separation task. In this paper, we present a novel-separation scheme that utilizes the complete auto-encoder architecture for separation. We do so by using the following setup for separation. Given an input spectrogram $\mathbf{X}$, the auto-encoder produces an approximation that models the spectrogram in terms of its weights. We will denote to this approximation as,
\begin{equation}
    \hat{\mathbf{X}} = Ae(\mathbf{X}|\theta)
    \label{eq:separation_ae}
\end{equation}
Here, $\theta$ denotes the weights (parameters) of the auto-encoder. Thus, given the trained auto-encoders for sources, i.e., given $\theta_{1}$ and $\theta_{2}$, the goal of separation is to identify suitable spectrograms $\mathbf{X}_{1}$ and $\mathbf{X}_{2}$ such that,
\begin{equation}
    \mathbf{X}_{m} = Ae(\mathbf{X}_{1}|\theta_{1}) + Ae(\mathbf{X}_{2}|\theta_{2}) 
    \label{eq:separation}
\end{equation}
Here, $\mathbf{X}_{m}$ represents the spectrogram of the mixture. In other words, the separation procedure attempts to estimate the mixture spectrogram as the sum of spectrograms of the individual sources. This assumption of additivity of spectrograms is similar to the separation procedure used in~\cite{}(**NMF, Icassp**). However, we directly estimate the source spectrograms without estimating the latent representation. We train the network defined by~(\ref{eq:separation}) for an appropriate input instead of training for the weights of the network. As before, we minimize the KL divergence between mixture spectrogram $\mathbf{X}_{m}$ and its approximation~$\mathbf{X}_{1} + \mathbf{X}_{2}$. Conceptually, this problem is not different to training a neural network. The equivalence can be seen by applying a transposition to the auto-encoder definitions in (\ref{eq:nmfcnncnn}). This also allows a generalized separation procedure to be applied even when the underlying architectures of the auto-encoders are changed.
